\documentclass{article}


% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024


% ready for submission
%\usepackage{neurips_2024}


% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2024}


% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage[final]{neurips_2024}


% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2024}


\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors


\title{Project Title: Exploring Representation Learning with Simple Images of Shapes}


% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.


\author{%
  Kevin Lewis\\
  \texttt{lewis.3164@osu.edu} \\
  \And
  Michael Lin\\
  \texttt{lin.3976@osu.edu} \\
    \And
  Shashank Raghuraj\\
  \texttt{raghuraj.2@osu.edu} \\
}


\begin{document}


\maketitle



\section{Project Type}
Reproducing examples in the textbook

\section{Project Introduction}
{\color{black}
In this project, our group will explore representation learning with simple images of shapes with autoencoders and contrastive learning. Representation learning allows us to automatically discover features from raw data.
In this project, we will replicate the textbook examples on autoencoders and contrastive learning, as seen here: 
\url{https://visionbook.mit.edu/representation_learning.html#experiment-do-autoencoders-learn-useful-representations}
and 
\url{https://visionbook.mit.edu/representation_learning.html#sec-representation_learning-expt_designing_embeddings_with_contrastive_learning}
} 


\section{Project Motivation}
% {\color{red}
% Please provide an motivation about why you choose to work on this project. It may be because the problem is timely, challenging, interesting, or because you want to explore it to strengthen your computer vision skills and understanding.
% } 

This project is interesting for a couple of reasons. First, the learning model does not seem too difficult itself, so we may be able to create both models without using any machine learning libraries.  This will be a useful learning tool for understanding more about how these models work behind the scenes.  Also, exploring the different things that are possible with the encoded representations (for example, animating a transition between different shapes with autoencoders).

\section{Project Plan}
\subsection{Baseline approach}
% {\color{red}
% Please specify the first (i.e., baseline) computer vision technique or algorithm you will try for your project and why? What are the potential limitations of this approach? If it is a neural network, please specify its architecture, such as Vision Transformer (ViT), Residual neural network (ResNet), etc.
% } 

We will use a simple fully connected neural network as a baseline approach for our project. This is because it should be enough for our purposes, and it is simple enough that we will be able to build it ground up.  One potential limitation of this is that using a fully connected network will limit the sizes of the hidden layers.  If they were too big, the model may be too big and slow to train effectively.

\subsection{Advanced approach}
% {\color{red}
% Please specify the computer vision techniques or algorithms you will develop and/or apply beyond the baseline approach? Why do you think these advanced approaches will lead to better performance or resolve the aforementioned limitations?
% } 

Beyond this, we may add some convolutional layers to see how they impact the performance and accuracy of the model. This will allow us to have a few layers without needing to decrease layer size.  We may also add some skip connections in these early layers to emulate a ResNet, in order to explore how this chance affects accuracy.

\subsection{Validation plan}
% {\color{red}
% How will you verify the effectiveness of your approach? This can be datasets and/or specific evaluation metrics. If there are training and test data, please specify the size, for example, 10,000 images across 10 categories. Do you have access to the data?
% }

We will have a dataset of shapes with different colors, sizes, and rotations to test our network on.  We do not currently have access to this dataset, but it will be easy to generate a dataset like it.  The textbook used 64,000 images, so we will look to generate a datasest of similar size.

Similarly to the textbook, we will do some nearest-neighbor search to validate our models.

\subsection{Computational resources}
% {\color{red}
% What computational resources do you plan to use?
% }

This model will be simple enough to run on our own computers.

\subsection{Library, Tool, etc.}
% {\color{red}
% What will be the programming language? What will be the library or tool you plan to use (e.g., PyTorch, scikit-image)?
% }
We will write this in Python.  The main library we use will likely be either Numpy or CuPy for fast matrix operations, though we will likely write the actual neural network code ourselves.  We may also use PyTorch if we choose to go with a machine learning library.

\subsection{Estimated baseline algorithm runtime}
% {\color{red}
% Please give an estimate of the baseline algorithm's training and/or test time.
% }

We are generating the dataset ourself, and since the dataset the texbook used is not too big, we can tune the size of our images so that one epoch worth of training takes at most a couple minutes.

\section{Workload}
{\color{black}
The workload for this project will consists of reading through examples to understand the prupose of autoencoders, and focus on the mathematical foundations they're used with. Next, we'll experiment with simple neural network implementations before diving into the primary models. The core workload will be the implementation of the model in which we will start by building a baseline fully connected network from scratch. We will spend the remainder of our time conducting experiments to evaluate the performance, accuracy, and representation quality.}


\section{Ideal result}
{\color{black}
If everything is successful, the autoencoder model should be able to reconstruct shapes from their compressed form. The contrastive model should produce embeddings that shows similarities between different images. The outcome will not only be a replica of the textbook experiments but also be well optimized
}

\subsection{Insights}
{\color{black}
We'll first understand how autoencoders learn compressed representations and how well they preserve essential features during the reconstruction of the object. We'll also learn how to map similar images closer together in the latent space. This project will also help us learn how to build a neural network from scratch.
}

\section{Potential Risk}
% {\color{red}
% Any potential risks that you may not complete your final project or your approach may not work? Please envision what may prevent you from completing the project. These may include your designed algorithm will not work as you expect, insufficient computational resources, etc. 
% }
It is possible we will have difficulty doing the calculus to get the project working. If this is too much of a problem for us, we can switch to using PyTorch, which should handle this for us.  

\section{Duplication Statement}
% {\color{red}
% If you choose a self-defined project, it is okay if it is related to your own research project (e.g., at your lab). However, you MUST write a short paragraph clearly articulating the difference between your final project and your own research project. More specifically, it is NOT allowed if you bring what you already planned to do in your own research project (before taking this class) as the final project.

% If you plan to reproduce an existing algorithm, please clearly specify whether there are accessible code/implementation online. If so, please clearly state how your project will not just be copying the code and rerunning it.
% }

There are surely autodencoders and contrastive learning models online, although not on the dataset we plan on using.  We will not use these models, and we will have to create our own transformations for contrastive learning.

%%%% Reference
\bibliographystyle{plain}
\bibliography{ref}


\end{document}